\documentclass[11pt,letterpaper]{article}

% =========================
% PACKAGES
% =========================
\usepackage[margin=0.5in]{geometry}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage[hidelinks]{hyperref}
\usepackage{fancyhdr}
\usepackage{tabularx}
\usepackage{array}
\usepackage{xcolor}
\usepackage{multicol}
\usepackage{ragged2e}

% =========================
% PAGE STYLE
% =========================
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\setlength{\parindent}{0pt}
\setlength{\tabcolsep}{0pt}

% =========================
% SECTION FORMATTING
% =========================
\titleformat{\section}
{\large\bfseries}
{}
{0em}
{}
[\color{black}\titlerule]

\titlespacing{\section}
{0pt}{8pt}{4pt}

% =========================
% CUSTOM COMMANDS
% =========================

\newcommand{\resumeSubheading}[4]{
\vspace{2pt}
\begin{tabularx}{\textwidth}{X r}
\textbf{#1} & #2 \\
\textit{#3} & \textit{#4}
\end{tabularx}
\vspace{-3pt}
}

\newcommand{\resumeProjectHeading}[4]{
\vspace{2pt}
\begin{tabularx}{\textwidth}{X r}
\textbf{#1} & #2 \\
\textit{#3} & #4
\end{tabularx}
\vspace{-3pt}
}

\newcommand{\resumeItem}[1]{
\item #1
}

\newcommand{\resumeSubHeadingListStart}{
\begin{itemize}[leftmargin=0.15in,label={}]
}

\newcommand{\resumeSubHeadingListEnd}{
\end{itemize}
}

\newcommand{\resumeItemListStart}{
\begin{itemize}[leftmargin=0.22in,itemsep=1pt,topsep=2pt]
}

\newcommand{\resumeItemListEnd}{
\end{itemize}
}

% =========================
% DOCUMENT START
% =========================

\begin{document}

% =========================
% HEADER
% =========================

\begin{center}
{\Huge \textbf{Dilip Sahu}} \\[4pt]

Aizawl, Mizoram (Current) \textbar\ Jaipur, Rajasthan (Home) \\[6pt]

+91-8233958847 \textbar\
\href{mailto:dilipsahu01022006@gmail.com}{dilipsahu01022006@gmail.com}
\textbar\
\href{https://www.linkedin.com/in/dilip-sahu-7b8789324/}{LinkedIn}
\textbar\
\href{https://github.com/Dilipsahu01}{GitHub}
\textbar\
\href{https://leetcode.com/}{LeetCode}
\textbar\
\href{https://dilipsahu01.github.io/}{Portfolio}
\end{center}

% =========================
% PROFESSIONAL SUMMARY
% =========================

\section{PROFESSIONAL SUMMARY}

2nd-year B.Tech Computer Science and Engineering student at National Institute of Technology Mizoram specializing in Scalable Distributed Systems, Artificial Intelligence Development, Internet of Things Systems, and Backend Engineering. National Level Qualifier (Top 200) in the Department of Telecommunications Government of India 5G Innovation Hackathon and multi-time contest winner with a strong foundation in Data Structures, Time-Series Analysis, and Large-Scale System Architecture. Expert in architecting full-stack applications with Third Normal Form normalized databases, real-time telemetry pipelines, and predictive Machine Learning models.

% =========================
% EDUCATION
% =========================

\section{EDUCATION}

\resumeSubheading
{National Institute of Technology (NIT), Mizoram}
{Expected May 2028}
{Bachelor of Technology in Computer Science and Engineering}
{Aizawl, India}

\resumeItemListStart
\resumeItem{CGPA: 7.98 / 10.00}
\resumeItem{Recent SGPA: 8.75}
\resumeItem{Relevant Coursework: Data Structures \& Algorithms, Object-Oriented Programming, System Programming, Computer Organization \& Architecture, Operating Systems, Database Management Systems, Linear Algebra, Probability \& Statistics}
\resumeItemListEnd

% =========================
% TECHNICAL SKILLS
% =========================

\section{TECHNICAL SKILLS}

\textbf{Programming Languages:}
C, C++, Java, Python, TypeScript, SQL, Bash

\vspace{2pt}

\textbf{Computer Science Foundations \& System Architecture:}
Data Structures \& Algorithms, Object-Oriented Programming, Model-View-Controller Architecture, Role-Based Access Control, Algorithm Complexity Analysis, Custom Neural Networks, Distributed Systems

\vspace{2pt}

\textbf{Artificial Intelligence, Machine Learning \& Data Science:}
Large Language Models (LLMs), Retrieval-Augmented Generation (RAG), Natural Language Processing (NLP), Meta Prophet, Random Forest, Claude AI, Whisper.cpp, SciPy, scikit-learn, Statsmodels, Pandas, NumPy

\vspace{2pt}

\textbf{Edge Artificial Intelligence \& Model Optimization:}
Model Quantization (1-bit, 4-bit, 8-bit), INT8 Quantization, On-Device Inference, Binary Neural Networks (BNNs), Edge AI

\vspace{2pt}

\textbf{Web Development \& Backend Engineering:}
FastAPI, Flask, Django, REST APIs, Next.js, React 18, Tailwind CSS, HTML5, CSS3, JavaScript, Apache Kafka, Flask-JWT-Extended, Jinja2

\vspace{2pt}

\textbf{Database Technologies:}
MySQL (3NF/BCNF), SQLAlchemy, ChromaDB, Supabase, H2 Database

\vspace{2pt}

\textbf{Cybersecurity \& Cryptography:}
Post-Quantum Cryptography (PQC), Lattice-Based Cryptography, AES, DES, RSA, SHA-256, Public Key Infrastructure (PKI), Side-Channel Attack Mitigation, Firmware Security

\vspace{2pt}

\textbf{Embedded Systems \& Internet of Things:}
ESP32 (C++), GSM/GPRS (SIM800L), NEO-6M GPS, UART Protocols, Wi-Fi Communication Protocols, Telemetry Failover Systems, Circular Buffers, Arduino

\vspace{2pt}

\textbf{Developer Tools \& Platforms:}
Distributed Version Control Systems, Source Code Management Workflows, Containerization Platforms, Reverse Proxy Servers, Linux (Ubuntu), Leaflet.js, Postman API Platform, PM2 Process Manager, Figma, Canva

% =========================
% EXPERIENCE, HACKATHONS
% =========================

\section{EXPERIENCE, HACKATHONS \& SIMULATIONS}

\resumeSubheading
{5G Innovation Hackathon 2026 | Department of Telecommunications (DoT), Government of India}
{May 2026 -- Present}
{National Level Qualifier (Top 200)}
{}

\resumeItemListStart
\resumeItem{Selected for the National Level Presentation-Based Evaluation after successfully clearing the highly competitive Zonal and State evaluation rounds.}
\resumeItem{Currently ranked within the top 200 teams nationwide, actively competing for a top 50 placement.}
\resumeItemListEnd

\resumeSubheading
{Google Cloud Rapid Agent Hackathon}
{May 2026}
{AI Developer (Multi-Agent Systems \& Google Cloud Platform)}
{Remote}

\resumeItemListStart
\resumeItem{Developing a ``Crisis Response Agent'' focusing on multi-step AI agents capable of taking real-world actions.}
\resumeItem{Utilizing Google Cloud Platform (GCP) and Retrieval-Augmented Generation approaches for dynamic data retrieval and rapid response generation.}
\resumeItemListEnd

\resumeSubheading
{JPMorgan Chase \& Co. (Forage)}
{March 2026}
{Software Engineering Virtual Experience}
{Virtual}

\resumeItemListStart
\resumeItem{Engineered a real-time data visualizer using Apache Kafka and H2 Database, achieving sub-10ms response latency for high-frequency live data ingestion.}
\resumeItem{Completed tasks in Project Setup, Kafka Integration, H2 Integration, REST API Integration, and REST API Controller Design.}
\resumeItemListEnd

\resumeSubheading
{OpenCV (Intel-led)}
{September 2025}
{Technical Contributor (C++)}
{Remote}

\resumeItemListStart
\resumeItem{Diagnosed a floating-point precision anomaly in OpenCV's Windows build caused by MinGW FPU extended-precision behavior.}
\resumeItem{Traced the root cause in C++ and submitted a validated Pull Request implementing strict double-precision semantics.}
\resumeItemListEnd

% =========================
% PART 1 END
% =========================

% CONTINUE WITH PART 2
% SECTION: KEY PROJECTS
% =========================
% KEY PROJECTS
% =========================

\section{KEY PROJECTS}

\resumeProjectHeading
{File Compressor Pro}
{Completed}
{HTML5, CSS3, Vanilla JS, Web Workers, PDF.js, jsPDF, heic2any}
{\href{https://digitalheroesco.com/}{built for DigitalHeros}}

\resumeItemListStart
\resumeItem{Engineered a high-performance, privacy-first, 100\% client-side file compression web application designed to run entirely in the browser using Web Workers for parallel processing.}
\resumeItem{Implemented hardware-aware concurrency control and asynchronous task queue to optimize resource utilization and prevent browser memory spikes/crashes during batch processes.}
\resumeItem{Designed a smart two-pass PDF rendering/downscaling engine utilizing PDF.js and jsPDF that decouples layout from resolution, preserves legibility, and handles transparencies cleanly.}
\resumeItemListEnd

\vspace{4pt}

\resumeProjectHeading
{CyberGuard-AI: On-Device Scam Call Detector}
{Apr 2026 -- Present}
{Python, Whisper.cpp, LoRA, GGUF, llama.cpp}
{\href{https://github.com/your-github-link}{GitHub} \textbar\ \href{https://your-live-demo-link.com}{Live Demo}}

\resumeItemListStart
\resumeItem{Building a privacy-first, fully on-device AI scam call detector for India with zero cloud dependency; targeting low-end Android phones and laptops (\textasciitilde480 MB footprint).}
\resumeItem{Pipeline: live microphone audio $\rightarrow$ 16-parameter extraction $\rightarrow$ Whisper.cpp tiny speech-to-text $\rightarrow$ language-aware LoRA adapter $\rightarrow$ conversation arc tracker $\rightarrow$ Q4 GGUF LLM classifier (SAFE / SUSPICIOUS / SCAM).}
\resumeItem{Generated 1,398 high-quality training samples using a 7B Large Language Model on Kaggle; supplemented with Indian High Court case records and YouTube fake call recordings.}
\resumeItemListEnd

\vspace{4pt}

\resumeProjectHeading
{TourAI: Smart Travel Planner}
{Apr 2026}
{FastAPI, React 18, Prophet, Random Forest, Claude 4, Containerization Platform}
{\href{https://github.com/your-github-link}{GitHub} \textbar\ \href{https://your-live-demo-link.com}{Live Demo}}

\resumeItemListStart
\resumeItem{Won 1st Place at ARTHKRITI AI Hackathon, co-developing the project alongside team members Naman Shukla and Anshul Singh.}
\resumeItem{Orchestrated a 6-module Machine Learning pipeline integrating cost forecasting via Meta Prophet (trained on a 15-year Mizoram dataset) and financial feasibility checks via Random Forest (100 estimators).}
\resumeItem{Integrated a Retrieval-Augmented Generation pipeline over local destination knowledge bases to provide accurate, grounded guidance via Claude Sonnet 4.}
\resumeItem{Deployed a full container orchestration stack with MySQL, FastAPI, and reverse-proxy-served React frontend, including a gamified milestone savings reward system.}
\resumeItemListEnd

\vspace{4pt}

\resumeProjectHeading
{IoT Campus Bus Tracker}
{Feb 2026 -- Present}
{ESP32, C++, SIM800L, NEO-6M GPS, Flask, Leaflet.js}
{\href{https://github.com/your-github-link}{GitHub} \textbar\ \href{https://your-live-demo-link.com}{Live Demo}}

\resumeItemListStart
\resumeItem{Architected an end-to-end telemetry system utilizing a 240MHz ESP32 with dual-network failover (GSM/Wi-Fi) and a 60-second timeout mechanism.}
\resumeItem{ESP32 Lead for a 5-member team; solely responsible for all C++ firmware, implementing a 10-packet circular buffer to guarantee data continuity during network outages.}
\resumeItem{Engineered hardware initialization logic to correctly account for the NEO-6M GPS module's hardware quirk, ensuring systems accounted for the delayed LED satellite fix signal.}
\resumeItem{Generated a 1-second telemetry cycle with approximately 2.5-meter GPS accuracy, transmitting JSON payloads containing latitude/longitude, speed in km/h, satellite count, HDOP, fix status, network type, and ISO-8601 UTC timestamp to a live Leaflet.js dashboard.}
\resumeItemListEnd

\vspace{4pt}

\resumeProjectHeading
{Custom Cryptographic Engine \& Firmware Security}
{Completed}
{C, Python, AES, DES, RSA}
{\href{https://github.com/your-github-link}{GitHub} \textbar\ \href{https://your-live-demo-link.com}{Live Demo}}

\resumeItemListStart
\resumeItem{Engineered custom implementations of AES, DES, and RSA from scratch, demonstrating deep understanding of core cryptographic primitives and bit-level data manipulation.}
\resumeItem{Developed secure hashing utilities (SHA-256) and digital signature verification protocols utilizing standard libraries for mathematical accuracy.}
\resumeItem{Designed firmware architectures resilient to side-channel attacks and fault injection techniques for embedded microcontrollers.}
\resumeItemListEnd

\vspace{4pt}

\resumeProjectHeading
{Full-Stack AI Document Intelligence Platform}
{Completed}
{Next.js, Django REST Framework, MySQL, ChromaDB, LM Studio}
{\href{https://github.com/your-github-link}{GitHub} \textbar\ \href{https://your-live-demo-link.com}{Live Demo}}

\resumeItemListStart
\resumeItem{Built a Retrieval-Augmented Generation ecosystem utilizing an automated Selenium scraper to ingest and normalize unstructured book data.}
\resumeItem{Implemented a dual-database architecture (MySQL for relational metadata, ChromaDB for vector embeddings) and local Large Language Models via LM Studio for zero-cloud, privacy-centric AI generation.}
\resumeItemListEnd

\vspace{4pt}

\resumeProjectHeading
{Senior Care Portal}
{Jan 2026 -- May 2026}
{Python, Flask, MySQL (3NF), SQLAlchemy, JWT, JavaScript}
{\href{https://github.com/your-github-link}{GitHub} \textbar\ \href{https://your-live-demo-link.com}{Live Demo}}

\resumeItemListStart
\resumeItem{Designed a 3NF-compliant 15-table relational schema to manage room capacities and full resident lifecycles (admit, transfer, discharge, re-admit).}
\resumeItem{Implemented database-level Role-Based Access Control (RBAC) via JWT tokens with stateless audit logging for medical data integrity.}
\resumeItemListEnd

\vspace{4pt}

\resumeProjectHeading
{Hyperliquid × Bitcoin Sentiment Analysis}
{Completed}
{Python, Pandas, SciPy, scikit-learn}
{\href{https://github.com/your-github-link}{GitHub} \textbar\ \href{https://your-live-demo-link.com}{Live Demo}}

\resumeItemListStart
\resumeItem{Conducted quantitative analysis on approximately 211K Hyperliquid perpetual futures trades merged with the Bitcoin Fear \& Greed Index to study sentiment regimes.}
\resumeItem{Utilized Kruskal-Wallis, Mann-Whitney, Spearman/Pearson correlations, and VaR/CVaR statistical tests for a Primetrade.ai data science hiring assignment, outputting 17 auto-generated dark-theme charts and an auto-rendered Markdown report.}
\resumeItemListEnd

\vspace{4pt}

\resumeProjectHeading
{Java Digit Recognizer (Binary Neural Network)}
{Nov 2025 -- Jan 2026}
{Java, MNIST, Swing, MVC}
{\href{https://github.com/your-github-link}{GitHub} \textbar\ \href{https://your-live-demo-link.com}{Live Demo}}

\resumeItemListStart
\resumeItem{Built a 3-layer fully connected neural network (784 input nodes $\rightarrow$ 128 hidden nodes via Sigmoid $\rightarrow$ 10 output nodes via Softmax) entirely from scratch in Core Java (JDK 8+) with manual backpropagation.}
\resumeItem{Implemented binary input encoding, deriving a 1-bit quantization concept to normalize pixels to a 0/1 representation, achieving 96\% accuracy on MNIST test images.}
\resumeItemListEnd

\vspace{4pt}

\resumeProjectHeading
{Water Sort Studio Pro}
{Oct 2025}
{Java, Swing, DFS/BFS Algorithms}
{\href{https://github.com/your-github-link}{GitHub} \textbar\ \href{https://your-live-demo-link.com}{Live Demo}}

\resumeItemListStart
\resumeItem{Engineered a desktop puzzle solver using state-space graph search (DFS/BFS) capable of solving 16-bottle configurations in under 2 seconds via pruning and memoization.}
\resumeItemListEnd

% =========================
% PART 2 END
% =========================

% CONTINUE WITH PART 3
% Leadership, Research Interests, Achievements, Certifications, Memberships, End Document
% =========================
% LEADERSHIP & EXTRACURRICULAR ACTIVITIES
% =========================

\section{LEADERSHIP \& EXTRACURRICULAR ACTIVITIES}

\resumeSubheading
{Coding \& AI Club, National Institute of Technology Mizoram}
{Feb 2025 -- Present}
{Team Head (Website Development \& Social Media)}
{Aizawl, Mizoram}

\resumeItemListStart
\resumeItem{Lead the official club website development utilizing HTML, CSS, JavaScript, and MySQL while conducting weekly Data Structures \& Algorithms and Distributed Version Control System workshops for more than 50 students.}
\resumeItem{Founded the ``Zero to Hero'' mentorship initiative focused on accelerating technical growth among junior students.}
\resumeItem{Coordinated alongside Club President Rishabh Shukla and Faculty Mentor Dr. Tanmay Kumar to execute the SAMYUGA technical event, managing results and operations for the Trishul, Chakravyuh, and Prarambh competitions.}
\resumeItemListEnd

\vspace{4pt}

\resumeSubheading
{TechnoX 2025--26}
{2026}
{Discord Server Handler \& Coordinator}
{}

\resumeItemListStart
\resumeItem{Managed the official Discord server for the university technical festival, ensuring seamless communication, participant engagement, community moderation, and technical support throughout event operations.}
\resumeItemListEnd

% =========================
% ACTIVE MEMBERSHIPS
% =========================

\section{PROFESSIONAL MEMBERSHIPS}

\resumeItemListStart
\resumeItem{\textbf{Arthkriti (Business \& Finance Club)} — Active Member (2025 -- Present)}
\resumeItem{\textbf{GDG on Campus (Google Developer Groups)} — Active Member (2025 -- Present)}
\resumeItem{\textbf{Think India} — Active Member (2025 -- Present); engaged in policy research and civic engagement initiatives}
\resumeItem{\textbf{JAD Foundation} — Active Member (2025 -- Present); involved in NGO activities and community outreach programs}
\resumeItemListEnd

% =========================
% RESEARCH INTERESTS
% =========================

\section{RESEARCH INTERESTS}

\resumeItemListStart

\resumeItem{\textbf{Post-Quantum Cryptography (PQC):} Researching lattice-based mathematical frameworks and cryptographic constructions designed to resist attacks from large-scale quantum computers.}

\resumeItem{\textbf{Model Quantization \& Edge Artificial Intelligence:} Actively exploring Binary Neural Networks (BinaryConnect, XNOR-Net), 1-bit Large Language Models (BitNet, BitNet b1.58), ultra-low-bit inference systems, and privacy-preserving on-device Artificial Intelligence deployment strategies for resource-constrained hardware.}

\resumeItem{\textbf{Distributed Systems \& Large-Scale Backend Engineering:} Interested in scalable service architectures, fault-tolerant infrastructure, telemetry pipelines, event-driven systems, and real-time data processing ecosystems.}

\resumeItem{\textbf{Embedded Artificial Intelligence \& Internet of Things:} Exploring efficient deployment of Machine Learning models on edge devices, microcontrollers, and low-resource computing environments.}

\resumeItemListEnd

% =========================
% ACHIEVEMENTS & AWARDS
% =========================

\section{ACHIEVEMENTS \& AWARDS}

\textbf{Hackathons \& Competitive Events}

\resumeItemListStart

\resumeItem{1st Place — ARTHKRITI Artificial Intelligence Hackathon.}

\resumeItem{National Level Qualifier (Top 200 Teams Nationwide) — Department of Telecommunications (DoT) Government of India 5G Innovation Hackathon 2026.}

\resumeItem{1st Place — Coding \& AI Club Data Structures \& Algorithms Challenge (2025).}

\resumeItem{1st Place — Coding \& AI Club Logo Design Competition.}

\resumeItemListEnd

\vspace{2pt}

\textbf{Ideathons, Innovation Challenges \& Strategic Pitches}

\resumeItemListStart

\resumeItem{3rd Position — Mathematics \& Computing Society Ideathon.}

\resumeItem{2nd Position — Sovereign Summit. Conceptualized and pitched ``The Abyssal Grid,'' a strategic roadmap for a Sovereign Blue Economy utilizing autonomous marine infrastructure, Ocean Thermal Energy Conversion (OTEC) plants, and next-generation maritime systems.}

\resumeItemListEnd

\vspace{2pt}

\textbf{Kaggle \& Data Science Competitions}

\resumeItemListStart

\resumeItem{Global Rank 1288 (Top 48\% among 2700+ teams) in the Deep Past Challenge (Akkadian-to-English Natural Language Processing Translation Competition). Achieved a score of 33.3263 against the competition-leading score of 42.8957.}

\resumeItemListEnd

\vspace{2pt}

\textbf{Competitive Programming}

\resumeItemListStart

\resumeItem{Solved 400+ problems across LeetCode and Striver's A2Z Problem Sheets.}

\resumeItem{Maintained an active 107-day LeetCode streak.}

\resumeItem{Maintained an active 45-day GeeksforGeeks Problem of the Day streak.}

\resumeItemListEnd

% =========================
% CERTIFICATIONS
% =========================

\section{CERTIFICATIONS}

\resumeItemListStart

\resumeItem{Fundamentals of Cybersecurity — Zscaler Academy.}

\resumeItem{Fundamentals of Cybersecurity — SWAYAM NPTEL.}

\resumeItem{Deloitte Technology Job Simulation Program.}

\resumeItem{Intellectual Property (IP) Management Workshop — February 2026.}

\resumeItem{Scaler Training Program.}

\resumeItem{5-Day Drone Technology Workshop.}

\resumeItemListEnd

% =========================
% PARTICIPATION & ENGAGEMENT
% =========================

\section{PARTICIPATION \& ACADEMIC ENGAGEMENT}

\resumeItemListStart

\resumeItem{Participated in multiple Bureau of Indian Standards (BIS) National Level Quizzes including United Nations Day Quiz, Independence Day Quiz, IEC Standards Quiz, and World Environment Day Quiz.}

\resumeItem{Actively involved in university-level hackathons, innovation challenges, technical competitions, open-source initiatives, and Artificial Intelligence development activities.}

\resumeItemListEnd

% =========================
% DOCUMENT END
% =========================

\end{document}
